---
title: "Using crisprapex"
author: Aditya Bhagwat, Mette Bentsen, Johannes Graumann, Mario Looso, 
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{finding_flank_cas9s}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

crisprapex makes the design of CRISPR-APEX experiments easier.
Serving to the specifics of such experiments, these functions:

* Find a CRISPR **set** to target a genomic **set**.
* Allow to target the **flanks** of genomic loci rather than loci themselves.
* Are **efficient**, building on top of data.table and Biostrings.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "    #"
)
```
```{r setup}
library(magrittr)
library(crisprapex)
```


# Find cas9 sites that target TF binding flanks

```{r}
bedfile <- system.file('extdata/SRF_sites.bed', package = 'crisprapex')
bsgenome <- BSgenome.Mmusculus.UCSC.mm10::Mmusculus
cas9dt <- find_flankonly_cas9s(bedfile, bsgenome)
```

# Limit to sites with no offtargets
```{r}
      cas9seqdt <- cas9dt [ , .(n0 = .N), by = cas9seq ]                            # Scan flanks

```




# Focus on good quality cas9 sites 

```{r}
do_score <- assertive.reflection::is_unix() & 
            reticulate::py_module_available('azimuth')
if (do_score){
  reticulate::use_condaenv('default')
  cas9dt [ !is.na(cas9start), ontargetscore := score_cas9ranges(chr, cas9start, cas9end, strand, bsgenome) ]
  message('\tFilter for scores > 0.4')
  cas9dt %<>% extract( ontargetscore>0.4 )
  message(sprintf('\t\t%d cas9 seqs across %d ranges', cas9dt[, length(unique(cas9seq))], nrow(cas9dt)))
}
```

